Cyclone: Fault Tolerance Middleware for NVM Clusters

### 2016-02-13 paper/paper.tex
More updates to paper.

\section{Programming Model}
Cyclone provides users with a replicated and durable heap that can be
manipulated via a set of library routines. The library provides special
functions to allocate and free persistent memory as well as fat pointers that
wrap pointers to persistent memory to ensure that they do not change across
restarts - since the virtual memory ranges into which persistent memory is
mapped may change.

@@

\section{Replication}

>>>>

\section{Single Node Programming Model}
Cyclone provides users with a replicated and durable heap that can be
manipulated via a set of library routines. The library provides special
functions to allocate and free persistent memory as well as fat pointers that
wrap pointers to persistent memory to ensure that they do not change across
restarts - since the virtual memory ranges into which persistent memory is
mapped may change.

@@

\section{Replication}
Cyclone replicates the state of the NVM heap across multiple machines in a
cluster. Cyclone provides strong consistency - a majority of machines in the
quorum of replicas always agrees on the current state of the NVM heap. Unlike
traditional approaches to strong replication however, Cyclone is designed to
decouple replication from execution of user code modifying NVM. It does so by
encapsulating changes in to the NVM heap in RPC calls driven by client requests.
We chose this model since it closely reflects the use cases most suitable for
NVM-based clusters.

Cyclone replicates state \emph{by executing the same sequence of RPC calls on
  all members of the quorum}. Unlike traditional replication therefore Cyclone
does not replicate logs of changes to apply to replicas. Instead it replicates
the RPC call and its arguments, executing it on every replica. By executing the
same sequence of RPC calls on all machines we end up with the same
state. Cyclone uses RAFT under the hood for consensus on the contents of the log
of RPC calls - although any other similar state machine replication protocol
such as Paxos is also suitable. Cyclone provides the means to achieve exactly
once semantics to clients by numbering RPC calls and remembering the results of
the last executed RPC call.

\subsection{RPC Call Numbering}
Cyclone is preconfigured for a fixed (large) number of clients. All RPC calls
from a client are numbered and Cyclone remembers the results of the last RPC
call made by a client. A particular client can exist on any machine - cyclone
treats all instances of the client as a single client. Cyclone rejects RPC calls
that are not in sequence. Any RPC call is executed exactly once across the
quorum - regardless of failure and reconfigurations. This provides the means for
clients to easily achieve exactly once semantics, provided the client does not
crash. Cyclone can supply the last executed RPC call number and its result to
assist recovering and stateless clients.

\subsection{Asynchronous Replication}
By default, Cyclone decouples execution of the RPC call on any replica from its
replication. This is particularly beneficial to performance since it allows
overlapping the update of non-volatile state with the network I/O associated
with replication. Persisting RAFT log entries, making the upcall to the
application and executing the RPC call in a failure atomic transaction takes a
significant amount of time and therefore it makes sense to overlap this with the
round-trip delay on the network for replication. This is possible in Cyclone and
not other traditional replication systems since we use failure atomic
transactions that cna be rolled back. Replication can fail due to various
reasons such as failure of nodes, a network partition or a new leader getting
elected in the underlying replication protocol (RAFT). Since failure atomic
transactions can be rolled back, we are able to speculatively apply the changes
to NVM without waiting for replication to finish.

The runtime steps associated with asynchronous replication are shown in
Figure~\ref{fig:async_rep}.

\begin{figure}
{\scriptsize
\begin{verbatim}
Initiate replication of RPC call and arguments
TX_BEGIN {
  Execute RPC call and modify local NVM
  while(!replication_complete);
  if(replication_failed) {
    TX_ABORT;
  }
} TX_END
\end{verbatim}
}
\caption{Asynchronous Replication}
\label{fig:async_rep}
\end{figure}

\subsection{Synchronous Replication}

\section{Multi-node Programming Model}

\subsection{Example: Distributed Transactions}
